---
title: "WGU_Queue_Data"
author: "Justin Howard"
date: "November 18, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading Data

```{r exploring structure}
library(tswge) # a handy package for time series analysis
MSDA_data = read.csv('MSDAsubmissions10.1.17_11.18.19.csv', header = T)
head(MSDA_data)
str(MSDA_data)
head(MSDA_data$ACTIVITY_STATUS_DATE)
str(MSDA_data$ACTIVITY_STATUS_DATE)
```
## Data Structure
Data is not grouped by date.
```{r restructuring for time series analysis}
MSDA_data2 = MSDA_data[order(MSDA_data$ACTIVITY_STATUS_DATE),]
# standardizing time to avoid CST, CDT time difference errors
Sys.setenv(TZ = 'GMT')
# creating univariate series
MSDA_t = strptime(MSDA_data2$ACTIVITY_STATUS_DATE, "%Y-%m-%d %H:%M:%S")
```
## Analysis of time between Data Analytics submissions
```{r}
time.lapses = c()
for (k in 1:length(MSDA_t)-1) {
  diff = abs(as.numeric(difftime(MSDA_t[k], MSDA_t[k+1], units = 'days')))
  time.lapses = c(time.lapses, diff)
}
summary(time.lapses)
head(time.lapses)
sum(is.na(time.lapses))
length(time.lapses) #9212
```

```{r autocorrelation scructure}
plotts.sample.wge(time.lapses)
```

## Initial Observations
### Realization
There does appear to be a trend in the duration of time between MSDA submissions
### Auto-Correlation Structure
The ACF structure resembles white noise
### Spectral Density
The system frequency rapidly decends from zero, also indicating that this series is white noise
### Implications for Analysis
The assumption of serial correlation is dubious at best. Linear regression will probably yield better results in predicting the average tiem between submissions.

```{r Getting Counts}
MSDA_data2$Hourly <- as.character( round(MSDA_t , units = "hours" ) )

hour_counts = aggregate( MSDA_data2$ACTIVITY_STATUS_DATE , by = list(MSDA_data2$Hourly) , length )
head(hour_counts)
plot(hour_counts$x, type = 'l', main = "Hourly Queue Activity")

MSDA_data2$Daily <- as.character( round(MSDA_t , units = "days" ) )
daily_counts = aggregate( MSDA_data2$ACTIVITY_STATUS_DATE , by = list(MSDA_data2$Daily) , length )
plot(daily_counts$x, type = 'l', main = 'Daily Queue Activity')
abline(v=235, col = 'red')
abline(v=185, col = 'red')
```
A very unusual spike in the data. What happened around day 200? 

```{r Univariate Exploratory Analysis}
library(tswge)
plotts.sample.wge(daily_counts$x)
```
Prominent spectral density peaks 
0, .08, .14, .28
Autocorrelation structure 
Significant lag = 7 --> weekly season
```{r Univariate model identification}
daily_q = daily_counts$x
s7 = artrans.wge(daily_q, phi.tr = c(rep(0,6),1))
```
An s=7 seasonal model seems to fit nicely. We will model the residuals
```{r Model ID con}
aic5.wge(s7, p = 0:7, q = 0:3)
```

AIC5 suggests an ARMA(5,2), the max allowed. We are missing something.
```{r yearly season}
s364 = artrans.wge(daily_q, phi.tr = c(rep(0,363),1))
aic5.wge(s364)
```
Trying AIC suggestion of s364 residuals.
```{r}
ARMA1.1 = est.arma.wge(s364, p = 1, q=1)

#sim1 = gen.aruma.wge(length(daily_q), s = 364, phi = ARMA1.1$phi, theta = ARMA1.1$theta, vara = ARMA1.1$avar)
```
generating realizations didn't work for some reason.
```{r forecast}
f1 = fore.aruma.wge(daily_q, phi = ARMA1.1$phi, ARMA1.1$theta, s = 364, n.ahead = 365, lastn = T, limits = F)
```
```{r ASE}
n = length(daily_q)
x = n-365
yr.test = daily_q[416:n]
plot(yr.test, type = 'l', main = 'ARIMA(1,0,1) s = 364 Forecast')
lines(f1$f, type = 'l', col = 'red')
mean((daily_q[416:n] - f1$f)^2)
sqrt((mean((daily_q[416:n] - f1$f)^2))/365)

f1residuals = f1$resid[366:780]
acf(f1residuals)
plot(f1residuals, type = 'l')
```
Although the residuals appear to be white noise, the ARIMA(1,0,1) model forecast doesn't capture the trend well. We can try differencing.

```{r DIFFERENCING}
Sys.setenv(TZ = 'GMT') # have to set to GMT otherwise positx provides different time zones
day1 = as.POSIXct(daily_counts$Group.1[1])
day1

last.d = as.POSIXct(daily_counts$Group.1[779])
last.d


dif = artrans.wge(daily_q, 1)
dif.s7 = artrans.wge(dif, phi.tr = c(rep(0,364),1))
aic5.wge(dif.s7, p=0:20)
phi = est.arma.wge(dif.s7, p=15, q=2)
f2 = fore.aruma.wge(daily_q, d = 1, s = 364, phi = phi$phi, theta = phi$theta, lastn = T, n.ahead = 364, limits = F)
```
```{r evaluating differenced model}
#plot(yr.test, type = 'l', main = 'ARIMA(1,0,1) s = 7 Forecast')
#lines(f2$f, type = 'l', col = 'red') 
mean((daily_q[416:n] - f2$f)^2) # 73.22, a little worse
sqrt((mean((daily_q[416:n] - f1$f)^2))/365)

f2residuals = f2$resid[366:780]
acf(f2residuals)
plot(f1residuals, type = 'l')
```
The differenced model appears to be white noise, and the ASE of the forecasts is slightly higher than the original model.
We are still failing to predict the rising trend well. We need to include time as a variable.
